---
title: "get Markov Equivalent Class"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obtaining the Markov equivalence class using pcalg

A quick document to check if my version of R and pcalg (or other dependencies) is causing a problem. First I create a non-collider 1->2->3 dag structure and a collider structure 1->2<-3. 

```{r dag}
library(pcalg)

p <- 3
# non-collider
(dag_NC_mat <- matrix(c(0,0,0,1,0,0,0,1,0),p,p))
dag_NC <- as(dag_NC_mat, "graphNEL")
# collider
(dag_C_mat <- matrix(c(0,0,0,1,0,1,0,0,0),p,p))
dag_C <- as(dag_C_mat, "graphNEL")
```


Lets look at the graphs:

```{r plotDagNC, echo=TRUE}
plot(dag_NC)
```

```{r plotDagC, echo=TRUE}
plot(dag_C)
```

Obtain cpdags corresponding to this generating dags, and then obtain all generated dags corresponding to cpdags. 

```{r cpdag}
cpdag_NC <- dag2cpdag(dag_NC)
cpdag_C <- dag2cpdag(dag_C)
```

Lets plot the cpdags. 

```{r plotcpagNC, echo=TRUE}
plot(cpdag_NC)
```

```{r plotcpdagC, echo=TRUE}
plot(cpdag_C)
```

Convert to matrices, obtain markov equivalent dags from each.

```{r cpdagMat}
# non-collider
cpdag_NC_Mat <- as(cpdag_NC, "matrix")
(dags_NC_Mat <- pdag2allDags(t(cpdag_NC_Mat))$dags)
dags_NC_aMat_1 <- matrix(dags_NC_Mat[1,],p,p,byrow=T)
dags_NC_aMat_2 <- matrix(dags_NC_Mat[2,],p,p,byrow=T)
dags_NC_aMat_3 <- matrix(dags_NC_Mat[3,],p,p,byrow=T)
# collider
cpdag_C_Mat <- as(cpdag_C, "matrix")
(dags_C_Mat <- pdag2allDags(t(cpdag_C_Mat))$dags)
dags_C_aMat_1 <- matrix(dags_C_Mat[1,],p,p,byrow=T)
```

As is specified in ?pdag2allDags pdag2allDags returns an adjacency of type "amat.cpdag"
which as is specified in ?amatType is coded so that mat[a,b]=0, mat[b,a]=1 implies a->b, 
so that "from" of arrow is in columns and "to" in rows. This means we have to transpose 
before converting to graph.

```{r dags_NC}
# non-collider
dags_NC_1 <- as(t(dags_NC_aMat_1),"graphNEL")
dags_NC_2 <- as(t(dags_NC_aMat_2),"graphNEL")
dags_NC_3 <- as(t(dags_NC_aMat_3),"graphNEL")
# collider
dags_C_1 <- as(t(dags_C_aMat_1),"graphNEL")
```

Lets plot non-collider generated dags. 

```{r plotDagNC_1, echo=TRUE}
plot(dags_NC_1)
```

```{r plotDagNC_2, echo=TRUE}
plot(dags_NC_2)
``` 

```{r plotDagNC_3, echo=TRUE}
plot(dags_NC_3)
```

This is fine! Now lets plot collider generated dags.

```{r plotDag_C1, echo=TRUE}
plot(dags_C_1)
```

The collider structure was converted into a non-collider structure in the process. 

This code, using the following packages, is returning a v-structure as one of the dags belonging to a Markov Equivalence class that was generated from a non-v-structure. 

```{r sessionInfo}
sessionInfo()
```
